---
title:  <center>FEMA Regional and Simulation Plots</center>
author: <center>Team</center>
date:   <center>`r Sys.Date()`</center>
output:
  html_document: 
    theme: united
    highlight: tango
    fig_height: 7
    toc: true
    number_sections: true
    toc_depth: 2
    toc_float: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = "~/2022 - FEMA")
```

```{r packages, include=FALSE}
library(MASS)
library(plotly)
library(ggplot2)
library(tidyverse)
library(data.table)
library(RColorBrewer)
```

# Mock Data EDA
## Load and clean
```{r mockData-load}
# Read in mock data file
fema <- read.csv(file = "./Attachment 6 mock_data.csv", header = T)

# Convert time features
fema$request_received_time <- as.POSIXct(fema$request_received_time,
                                       format="%Y-%m-%dT%H:%M:%S")
fema$deployment_start_time <- as.POSIXct(fema$deployment_start_time,
                                       format="%Y-%m-%dT%H:%M:%S")
fema$deployment_end_time   <- as.POSIXct(fema$deployment_end_time,
                                       format="%Y-%m-%dT%H:%M:%S")
```

## Generate additional features
```{r features-generate}
# Request interval time - all requests (request_time(n+1) - request_time(n))
fema <- fema %>%
  arrange(request_id) %>%
  mutate(interval_time = difftime(request_received_time,
                                  lag(request_received_time), units='days'))

# Request interval time - by responder
fema <- fema %>%
  arrange(responder_id, request_received_time) %>%
  group_by(responder_id) %>%
  mutate(int_time_resp = difftime(request_received_time,
                                  lag(request_received_time), units='days'))

# Fulfillment time (start_time - request_time)
fema$fulfill_time <- difftime(fema$deployment_start_time,
                            fema$request_received_time, units='days')

# Execution time (end_time - start_time)
fema$execution_time <- difftime(fema$deployment_end_time,
                              fema$deployment_start_time, units='days')

# Bench time - by responder (start_time(n+1) - end_time(n))
fema <- fema %>%
  arrange(responder_id, request_received_time) %>%
  group_by(responder_id) %>%
  mutate(bench_time_resp = difftime(deployment_start_time,
                                    lag(deployment_end_time), units='days'))

# Utilization rate by responder
fema <- fema %>%
  group_by(responder_id) %>%
  mutate(util_rate = sum(execution_time)/8760)

# Bench rate by responder
fema <- fema %>%
  group_by(responder_id) %>%
  mutate(bench_rate = sum(bench_time_resp, na.rm=T)/8760)
```

## Calculate summary statistics
```{r summary-stats}
# Summary statistics
cat("Summary Stats",
    "\n", "Mean fulfillment time:", mean(fema$fulfill_time), "days",
    "\n", "Mean execution time:", mean(fema$execution_time), "days",
    "\n", "Yearly employment rate:",
             sum(as.numeric(fema$execution_time))/
               (sum(as.numeric(fema$execution_time))+
                  sum(as.numeric(fema$bench_time_resp), na.rm = T)), "percent",
    "\n", "Yearly deployments: ", nrow(fema),
    "\n", "Team Size: ", length(unique(fema$responder_id)))
```

## Util and bench rates by month
```{r summ-util-bench}
# Extract month from start_time
fema$month <- months(fema$deployment_start_time)

# Responder execution time by month
summ <- fema %>%
  arrange(month,responder_id) %>%
  group_by(month,responder_id) %>%
  mutate(month_exec = sum(execution_time))

summ <- summ %>%
  group_by(month) %>%
  mutate(mean_exec_rate = mean(month_exec)/720)

# Responder bench time by month
summ <- summ %>%
  arrange(month,responder_id) %>%
  group_by(month,responder_id) %>%
  mutate(month_bench = sum(bench_time_resp, na.rm=T))

summ <- summ %>%
  group_by(month) %>%
  mutate(mean_bench_rate = mean(month_bench)/720)

summ <- summ[!duplicated(summ$month), ]
```

# Mock Data Plots
## Execution time by responder
```{r box-exec-time, fig.align="center", fig.width=5, fig.height=5, fig.cap="<b>Figure 1</b> Execution time by responder"}
boxplot(fema$execution_time ~ fema$responder_id, horizontal=TRUE,
        main="Responder Task Execution Times",
        ylab="Responder ID", 
        xlab="Days")
```

## Bench time by responder
```{r box-idle-time, fig.align="center", fig.width=5, fig.height=5, fig.cap="<b>Figure 2</b> Bench time by responder"}
boxplot(fema$bench_time_resp ~ fema$responder_id, horizontal=TRUE, 
        main="Responder Bench Time", 
        ylab="Responder ID", 
        xlab="Days")
```

## Utilization and bench rates
```{r bar-util-bench, fig.align="center", fig.width=5, fig.height=5, fig.cap="<b>Figure 3</b> Utilization and bench rates by month"}
summ$month <- as.character(summ$month)
summ$month <- factor(summ$month, levels=c("January","February","March",
                                          "April","May","June",
                                          "July","August","September",
                                          "October","November","December"))

ggplot(data=summ, aes(x=month)) +
  geom_bar(aes(y=mean_exec_rate), stat="identity", position ="identity", alpha=.3,
           fill='lightblue', color='lightblue4') +
  geom_bar(aes(y=mean_bench_rate), stat="identity", position="identity", alpha=.8,
           fill='pink', color='red') +
  ggtitle("Utilization and Bench Rates by Month") +
  labs(y="Rate", x="Responder ID")
```

## Time series utilization
```{r gantt-util-bench, fig.align="center", fig.width=5, fig.height=5, fig.cap="<b>Figure 4</b> Time series utilization by responder"}
ggplot(fema, aes(x=deployment_start_time, xend=deployment_end_time,
               y=responder_id, yend=responder_id)) +
  theme_bw() + 
  geom_segment(size=2) + 
  labs(title='Execution and Bench Times', x='Date', y='Responder ID')
```

## Interval time distribution
```{r hist-gap-time, fig.align="center", fig.width=5, fig.height=5, fig.cap="<b>Figure 5a</b> Distribution of request interval times"}
ggplot(fema, aes(x = as.numeric(fema$interval_time))) + 
  geom_histogram(colour = 4,
                 fill = "white",
                 lwd = 0.75,
                 linetype = 1,
                 bins = 20,
                 position = "identity") +
  ggtitle("Interval Time Between Requests") +
  labs(y = "Frequency", x = "Days")
```

## Interval time by date
```{r gap-by-date, fig.align="center", fig.width=5, fig.height=5, fig.cap="<b>Figure 5b</b> Request interval times by date"}
ggplot(fema, aes(x=request_received_time, y=as.numeric(interval_time))) +
  geom_line(size=0.7, colour = 4) + 
  ggtitle("Time Between Requests by Date") +
  labs(y="Days", x="Request Received Date")
```

## Fulfillment time distribution
```{r hist-fulfill-time, fig.align="center", fig.width=5, fig.height=5, fig.cap="<b>Figure 6a</b> Distribution of request fulfillment time"}
ggplot(fema, aes(x = as.numeric(fema$fulfill_time))) + 
  geom_histogram(colour = 4,
                 fill = "white",
                 lwd = 0.75,
                 linetype = 1,
                 bins = 20,
                 position = "identity") +
  ggtitle("Time to Fulfill Request") +
  labs(y = "Frequency", x = "Days")
```

## Fulfillment time by date
```{r fulfill-by-date, fig.align="center", fig.width=5, fig.height=5, fig.cap="<b>Figure 6b</b> Request fulfillments times by date"}
ggplot(fema, aes(x=request_received_time, y=as.numeric(fulfill_time))) +
  geom_line(size = 0.7, colour = 4) + 
  ggtitle("Time to Fulfill Request by Date") +
  labs(y="Days", x="Request Received Date")
```

## Interval by execution time
```{r scatter-gap-exec, fig.align="center", fig.width=5, fig.height=5, fig.cap="<b>Figure 7</b> Scatter plot of request interval time by execution time"}
ggplot(fema, aes(as.numeric(fema$interval_time), as.numeric(fema$execution_time))) +
  geom_point() +
  ggtitle("Interval Time by Execution Time") +
  labs(y = "Execution Time (Days)", x = "Interval Time (Days)")
```

# Simulation Data
## Source simulation functions
```{r load-functions}
source('./fema_code_challenge/simulation/deployment_simulation.r')
```

## Generate simulation stats
```{r generate-sim-data}
# Set x number of years worth of deployments (1)
deployments <- generate_deployments(1)

# Set range of team sizes (30:49)
proc_list <- list()

for (i in 30:49){
  sim <- process_deployments(i,deployments)
  sim$Team_Size <- i
  sim <- sim %>%
    arrange(responder_id, request_received_time) %>%
    group_by(responder_id) %>%
    mutate(idle_time = deployment_start - lag(deployment_end))
  proc_list[[i]] <- sim
}

# Combine data files for each team size and generate summary stats
sim_data = do.call(rbind, proc_list)

summary <- sim_data %>%
  group_by(Team_Size) %>%
  summarize(Total_Exec = sum(deployment_length)/24,
            Total_Fulfill = sum(deployment_delay)/24,
            Total_Idle = sum(idle_time, na.rm = T)/24) %>%
  mutate(Util_Rate = Total_Exec/(Total_Exec+Total_Idle))
```

# Simulation Data Plots
## Total fulfill time by team size
```{r fulfill-by-responders2, fig.align="center", fig.width=5, fig.height=5, fig.cap="<b>Figure 8</b> Time to fulfill request by team size"}
plot(x=summary$Team_Size, y=summary$Total_Fulfill,
     main = "Time to Fulfill Request by Team Size",
     xlab = "Team Size",
     ylab = "Days")
```

## Utilization rate by team size
```{r employ-by-responders, fig.align="center", fig.width=5, fig.height=5, fig.cap="<b>Figure 9</b> Employment rate of team members by team size"}
plot(x=summary$Team_Size, y=summary$Util_Rate,
     main = "Team Member Utilization Rate (Efficiency)",
     xlab = "Team Size",
     ylab = "Employment Rate")
```