---
title: "FEMA Code Challenge - Statistical Analysis"
author: "Nikolai Shaposhnikov"
date: "2022-08-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simulation Tests

In this session we test how the simulation code.

```{r}
library(ggplot2)
library(ggalt)

source('../simulation/deployment_simulation.r')
```

## Run simulations

```{r}
num_years = 5
responder_num = 31

sim <- simulate_deployments(5)
metrics <- sim$metrics
deployments = resp$data
processed_deployments <- process_deployments(responder_num = responder_num,deployments)
summary(processed_deployments)
```

Plot the deployment timeline:
```{r}
gg <- ggplot(processed_deployments, 
             aes(
               x=deployment_start, 
               xend=deployment_end, 
               y=responder_id, group=responder_id),xlab="Time, hrs") + 
        geom_dumbbell(color="#00ba38", 
                      size=1.5,
                      size_x = 0.0,
                      size_xend = 0.0) + 
        #scale_x_continuous(label=deployment_delay) + 
        labs(x='Time, Hrs', 
             y='Responder ID', 
             title="Deployments Chart", 
             subtitle="Simulated data", 
             caption="") +
        theme(plot.title = element_text(hjust=0.5, face="bold"),
              plot.background=element_rect(fill="#f7f7f7"),
              panel.background=element_rect(fill="#f7f7f7"),
              panel.grid.minor=element_blank(),
              panel.grid.major.y=element_blank(),
              panel.grid.major.x=element_line(),
              axis.ticks=element_blank(),
              legend.position="top",
              panel.border=element_blank())
plot(gg)
```
Calculate the number of deployments in the cue:
```{r}
num_of_deployments = length(processed_deployments$responder_id)

num_of_depl_received <- integer(num_of_deployments)
num_of_depl_started <- integer(num_of_deployments)
num_of_depl_ended <- integer(num_of_deployments)

for (i in (1:num_of_deployments)){
  t = processed_deployments$request_received_time[i]
  num_of_depl_received[i] <- sum(as.numeric(processed_deployments$request_received_time < t))
  num_of_depl_started[i] <- sum(as.numeric(processed_deployments$deployment_start < t))
  num_of_depl_ended[i] <- sum(as.numeric(processed_deployments$deployment_end < t))
}

depl_num <- num_of_depl_started - num_of_depl_ended
cue_num <- num_of_depl_received - num_of_depl_started
```


Plot the number of deployments in the cue:
```{r}
plot(processed_deployments$deployment_start,cue_num, type = 'l' , 
     ylab= 'Number of deployments', xlab = 'Time, Hrs', main = "Number of deployments in the cue")

```
```{r}
plot(processed_deployments$deployment_start,depl_num, type = 'l' , 
     ylab= 'Number of deployments', xlab = 'Time, Hrs', main = "Number of deployments in execution")

```
Calculate suggested number of responders.
The `cost_ratio` is the ratio of the cost of an hour of deployment fulfillment delay to the costs of an hour of a responder bench time:


```{r}
cost_ratio = 0.5
suggested_responder_num <- suggest_responder_num(resp,0.5)
suggested_responder_num
```



Run tests:

```{r}
library(testthat)
testthat::test_dir('../tests')
```